\begin{center}
\textbf{Abstract}
\end{center}

\red{could one say its the first application of a systematic aware optimization in a realistic particle physics search?}
\noindent This thesis presents the first application of a novel optimization technique, termed - \acf{neos} -, in a search for Higgs boson pairs decaying into four $b$-quarks with a boosted vector boson fusion topology at the ATLAS experiment. Particle physics research often involves complex, multi-stage analyses requiring iterative optimizations, typically reliant on signal-to-background ratios. While these conventional optimizations are linked to significance measures such as p-values, they cannot account for the critical role of systematic uncertainties when determining p-values, nor can they consider the complex interdependencies across the analysis pipeline. The \ac{neos} method revolutionizes this by optimizing directly on the p-value, considering all aspects of the analysis simultaneously.

This study shows that the \ac{neos} approach can effectively optimize any parameterizable method or parameter in a realistic particle physics search, demonstrated in this work by the optimization of event filtering and background estimation methods. It also introduces the concept of 'autoanalysis', where the algorithm autonomously determines the optimal analysis configuration for a given analysis strategy using only basic reconstructed physics objects. This innovation allows researchers to focus on developing analysis strategies rather than on manual optimization tasks.

The analysis uses the full Run 2 data from the ATLAS detector, incorporating the latest advancements in object reconstruction, such as Unified Flow Object large-radius jets and the GN2X version of the $X\rightarrow bb$-tagging algorithm. The \ac{neos} method has shown to reduce the expected cross-section limits by a factor of 2 over traditional methods that use the invariant mass of the Higgs pair system and a classifier neural network optimized for signal-to-background ratio. With \ac{neos} and these reconstruction advancements, expected constraints for the \ktwov coupling are set to \red{$(0.55 < \kappa_{2v}< 1.48)$} at the 95\% confidence level. This work establishes \ac{neos} as a robust tool for optimizing particle physics analyses, highlighting its potential for future research and discovery.

Furthermore, this thesis presents an improved strategy for identifying $b$-quarks in small-$R$ jets, crucial for detecting Higgs pairs in the four $b$-quark final state. This method, which uses muons from semileptonic hadron decays, improves background rejection by 40%.


% fall short in addressing 

% these conventional approaches often overlook systematic uncertainties and the intricate interdependencies throughout the analysis pipeline.


